---
title: "p8105_hw2_ll3407"
Name: Liping Lu
Uni: ll3407
output: github_document
date: "2022-10-02"
---

## Problem 1
```{r, eval=FALSE}
library(tidyverse)
NYC_transit = read_csv(file = './data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
janitor::clean_names() %>% 
  select(line,station_name,station_latitude,station_longitude,route1:route11, entrance_type, entry, vending, ada) %>%     mutate(entry=if_else(entry=="NO",FALSE,TRUE)) 

```

Description of dataset:  
* The original data 'NYC_Transit_Subway_Entrance_And_Exit_Data.csv' contains 32 variables, including Division, Line, Station_name, Station_latitude, Station_longitude, Routes, etc.  
* We only retain variables into a new data "NYC_transit", including line, station,name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
* There are 1,868 rows x 19 columns in data 'NYC_transit'. Those data are not tidy.

Stations distinct
```{r eval=FALSE}
  dplyr::distinct(NYC_transit, station_name, .keep_all = FALSE)
```
* There are 356 distinct stations.

```{r,eval=FALSE}
table(NYC_transit$ada)
```

* 468 stations are ADA compliant.

```{r,eval=FALSE}
table(NYC_transit$vending)
```

* The proportion of station entrances/exits without vending allow entrance is `r 100*(183/(183+1685))`%

```{r, eval=FALSE}
NYC_transit$route8<-as.character(NYC_clean$route8, c(1,4,5,23,4,20,NA))
NYC_transit$route9<-as.character(NYC_clean$route9, c(2,5,23,5,NA))
NYC_transit$route10<-as.character(NYC_clean$route10, c(3,23,NA))
NYC_transit$route11<-as.character(NYC_clean$route11, c(7,23,NA))

NYC_route =
tidyr::pivot_longer(
    NYC_transit,
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
    names_prefix = "route"
    )
view(NYC_route）
table(NYC_route$route_name)
```
* There are 273 distinct stations serve the A train.

```{r,eval=FALSE}
A_ada=filter(NYC_route,route_name =="A" & ada == "TRUE")
view(A_ada)
```
* Of the stations that serve the A train, there are 107 ADA compliant.

##Problem 2  
*Read and clean Mr.Trash Wheel dataset
```{r,eval=FALSE}
MTW_data = 
  readxl::read_excel('./data/Trash Wheel Collection Data.xlsx', sheet = 'Mr. Trash Wheel', range = "A2:N550") %>% 
  janitor::clean_names() %>% 
  tidyr::drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer((sports_balls),length=0)) %>% 
  mutate(type="Mr")

```

*Read and clean Mr.Trash Wheel dataset
```{r,eval=FALSE}
PTW_data = 
  readxl::read_excel('./data/Trash Wheel Collection Data.xlsx', sheet = 'Professor Trash Wheel', range = "A2:M97") %>% 
  janitor::clean_names() %>% 
  tidyr::drop_na(dumpster) %>% 
  mutate(type = "Professor") %>% 
  mutate(year = as.character(year))

   
```

* a single tidy data
```{r, eval=FALSE}
single_tidy = 
  bind_rows(MTW_data,PTW_data)

sum(PTW_data$weight_tons)

MTW2020=filter(MTW_data,year=='2020')
sum(MTW2020$sports_balls)
```
*_The combined data 'single_tidy' has 641 observations and 15 variables, including dumpster, month, year, date, weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_balls, homes_powered, and type._  

*_The total weight of trash collected by Professor Trash Wheel is 190.12 tons._   

*_The total number of sports balls collected by Mr. Trash Wheel in 2020 is 856._

## Problem 3
* Clean data pols-month.csv

```{r eval=FALSE}

library(tidyverse)
pols = read_csv(file = './data/Fivethirtyeight/pols-month.csv' )%>%   janitor::clean_names() %>% 
 separate(mon, sep="-", into = c("year","month","day" ))%>% 
  mutate(month= recode(month, "01" = "jan", "02" = "feb", "03" = "mar", "04" = "apr", "05" = "may", "06" = "jun", "07" = "jul", "08" = "aug", "09" = "sep", "10" = "oct", "11" = "nov", "12" = "dec")) %>% 
  pivot_longer(prez_gop|prez_dem,
               names_to = 'president',
               names_prefix = 'prez_') %>% 
  select(-day) %>% 
  mutate(date = paste(year, month, sep = "-"))
```

* Clean data snp.csv
```{r eval=FALSE}
library(lubridate)
snp = read_csv(file = './data/Fivethirtyeight/snp.csv')%>% 
  janitor::clean_names()%>%
  mutate(date= as.Date(date, "%m/%d/%y"))%>% 
  mutate(date=as.Date(ifelse(date > Sys.Date(), 
  format(date, "19%y-%m-%d"), 
  format(date)))) %>% 
  mutate(date = as.character(date)) %>% 
    separate(
  date, sep="-", into = c("year","month","day"),remove = FALSE) %>% 
  mutate(month = recode(month, "01" = "jan", "02" = "feb", "03" = "mar", "04" = "apr", "05" = "may", "06" = "jun", "07" = "jul", "08" = "aug", "09" = "sep", "10" = "oct", "11" = "nov", "12" = "dec")) %>%
  select(-day) %>% 
  select(-date) %>% 
  arrange(year, month) %>% 
  relocate(year,month) %>% 
  mutate(date = paste(year, month, sep = "-"))
```

* Tidy unemployment data
```{r,eval=FALSE}
unemployment = read_csv(file = './data/Fivethirtyeight/unemployment.csv')%>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage of unemployment") %>% 
    mutate(date = paste(year, month, sep = "-"))

```


```{r,eval=FALSE, include=FALSE}
pols_snp = left_join(pols, snp, by="date")
final = left_join(pols_snp, unemployment, by = "date")
arrange(final,year.x)
```


Description  
* The final data consists of three data, including pols (main data), SNP, and unemployment.  
* The Pols data contains 11 variables and 1644 observations, relating to the number of national politicians who are democratic or republican at any given time. The key valuables include year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, president, value, and date.  
* The snp data contains 787 observations and 4 variables relating to Standard & Poor's stock market index. The variables inlcude year, month,close, and date.  
* The unemployment data contains 816 observations and 4 variables (year, month, percentage of unemployment, and date) related to the percentage of unemployment.  
* We merge snp into pols, and then merge unemployment to the combined data, and have a final data.  
* The final data contains 1644 observations and 19 variables.  
* The range of year is 1947-2015.  
* The key variables include date, president, close, percentage of unemployment, etc.





