---
title: "p8105_hw2_ll3407"
output: github_document
date: "2022-10-02"
---

## Problem 1
Read and clean the data
```{r, echo=FALSE}
library(tidyverse)
NYC_transit = read_csv(file = './data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
janitor::clean_names() 
view(NYC_transit)
names(NYC_transit)
```

Retain line, station,name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance


```{r}
NYC_clean = select(NYC_transit,line,station_name,station_latitude,station_longitude,route1:route11, entrance_type, entry, vending, ada)

```

Convert entry variable from character to a logical variable
```{r}
NYC_clean$entry <-NYC_clean$entry %>% 
plyr::revalue(c("YES"=TRUE,"NO"=FALSE))
```
Description of dataset:  
* The original data 'NYC_Transit_Subway_Entrance_And_Exit_Data.csv' contains 32 variables, including `r names(NYC_transit)`  
* We clean the name of the dataframe, retain variables into the new dataframe 'NYC_clean' including line, station,name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance  
* There are 1,868 rows x 19 columns in data 'NYC_clean'. Those data are not tidy.

Stations distinct
```{r}
distinct(NYC_clean, station_name, .keep_all = FALSE)

```
* There are 356 distinct stations.

```{r}
table(NYC_clean$ada)
```

* 468 stations are ADA compliant.
```{r}
table(NYC_clean$vending)
```

* The proportion of station entrances/exits without vending allow entrance is `r 100*(183/(183+1685))`%
